What is happening on Twitter? A framework for student research projects with tweets

### 2020-09-07 Rmd/jse-special-issue-2020-october.Rmd
added LDA generative process math. notation.

Latent Dirichlet allocation is a statistical method for inferring latent (unobservable) topics (or themes)
from a large corpus (or collection) of documents [@blei2003latent].
We pretend that there's an imaginary process for creating documents in the corpus. For each document,
we choose a discrete distribution over topics. For example, some tweets from Memorial Day may 
refer to the holiday. This may constitute one topic in the corpus. 
Having chosen a distribution over topics, we then select document words by first 
drawing a topic from the distribution over topics, then drawing a word from the
chosen topic. 
The goal for latent Dirichlet allocation is to infer both the distribution over topics and 
the topics [@blei2003latent]. A topic, in this setting, is a distribution over 
the vocabulary (the collection of all words in a corpus).

Inference for latent Dirichlet allocation models is performed by either 
sampling from the posterior distribution or through variational methods. Researchers
have devised a variety of Gibbs sampling techniques for these models [@porteous2008fast]. 
Variational methods, while using approximations to the posterior distribution, offer the 
advantage of computational speed [@blei2017variational]. We used variational methods below.

>>>>

Latent Dirichlet allocation is a statistical method for inferring latent (unobservable) topics (or themes)
from a large corpus (or collection) of documents [@blei2003latent].
We pretend that there's an imaginary process for creating documents in the corpus. For each document,
we choose a discrete distribution over topics. For example, some tweets from Memorial Day may 
refer to the holiday. This may constitute one topic in the corpus. 
Having chosen a distribution over topics, we then select document words by first 
drawing a topic from the distribution over topics, then drawing a word from the
chosen topic.

In mathematical notation, we write the generative process assumed by LDA [@blei2003latent]:

1. Choose $N \sim \text{Poisson}(\xi)$  
1. Choose $\theta \sim \text{Dirichlet}(\alpha)$  
1. For each word, $w_n$ with $n = 1, ..., N$, 
  a. Choose a topic $z_n \sim \text{Multinomial}(\theta)$  
  b. Choose a word $w_n$ from $p(w_n | z_n, \beta)$, a multinomial probability

$\beta$ refers to the $k$ by $V$ matrix of topic-specific word probabilities, where $k$
is the number of topics and $V$ is the size of the vocabulary, *i.e.*, the number of unique words in the corpus.

The goal for latent Dirichlet allocation is to infer both the distribution over topics and 
the topics [@blei2003latent]. A topic, in this setting, is a distribution over 
the vocabulary (the collection of all words in a corpus).

Inference for latent Dirichlet allocation models is performed by either 
sampling from the posterior distribution or through variational methods. Researchers
have devised a variety of Gibbs sampling techniques for these models [@porteous2008fast]. 
Variational methods, while using approximations to the posterior distribution, offer the 
advantage of computational speed [@blei2017variational]. We used variational methods below in our case study.
